---
title: "sac_cc_districts"
output: github_document
---
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(sf)
library(tidygeocoder)
library(tidyverse)
library(ggmap)
setwd("~/coding/sac_cc_districts")
```

**Locating City Council District based on Address**

This is a notebook I created to locate city council districts given an address. 
It's meant to work with downloaded data from Action Network. 

This work was done for [Sac Kids First](https://sackidsfirst.org/). 

### Read in Council District KML Data 

This dataset is from [Sacramento Open Data](https://data.cityofsacramento.org). 
It's very well maintained! I'm using the `sf` package to read it in. 

```{r}
cc <- st_read("Council_Districts.kml")
cc[1]
```


### Load in Addresses, Convert to Coordinates (Geocoding)

Now we're loading in a bunch of random addresses I picked from around the Sacramento region.
It's in the same format from Action Network so it should work with the activist data.
On the second line, I use Regular Expressions (Regex) to remove apartment/unit numbers which 
can create problems.

```{r}
addresses <- read.csv('test.csv')

# (?i) - case insenstive match
addresses$address <- gsub("(\\#|(?i)apt|(?i)unit).*", "", 
                               addresses$address, perl=TRUE)
#remove commas & periods
addresses$address <- gsub("(\\,|\\.)", "", 
                               addresses$address, perl=TRUE)

addresses <- addresses %>% 
  mutate(full_address = paste(address, city, state, zip_code, sep=", ")) 
```

We then use `geocode` from `tinygeocoder`. It's a very handy function and easy to use. 
Here is the result. Note that some places are outside of the city limits. 

```{r message=FALSE, warning=FALSE}
addresses  <- tidygeocoder::geocode(addresses %>% select(name, full_address), full_address, method = "osm")
addresses 
```

### Find CC District based on Address

Now we use st_intersecs to find out which addresss is in which district. 

```{r}
pnts = addresses %>% select(long, lat)

address_sf <- do.call("st_sfc",c(lapply(1:nrow(pnts), 
function(i) {st_point(as.numeric(pnts[i, ]))}), list("crs" = 4326))) 

address_trans <- st_transform(address_sf, 2163) # apply transformation to pnts sf
cc_trans <- st_transform(cc[1], 2163)      # apply transformation to polygons sf

# intersect and extract state name
addresses$council_member <- apply(st_intersects(cc_trans, address_trans, sparse = FALSE), 2, 
               function(col) { 
                  cc_trans[which(col), ]$Name
               }) %>% as.character()

```


We create a small dataframe to match each council member to a district number. 

```{r}
districts = data.frame(council_member = c("Angelique Ashby", "Sean Loloee", 
                                          "Jeff Harris", "Katie Valenzuela", 
                                          "Jay Schenirer", "Eric Guerra",  
                                            "Rick Jennings II", "Mai Vang"), 
         cc_district = 1:8)
```

Now we left_join the data frame above and we have our result!
Note that the following places are outside of the city limits so has no council district:

* 99 Ranch (off of highway 99 - not in city limits)
* Kingdom Coffee Roasters (Folsom)
* Temple Coffee (Arden)

```{r message = FALSE}
addresses %>% select(name, full_address, council_member) %>% 
            left_join(districts)
```

### Plotting the Results

Here I download the basemap for the Sacramento region. 
```{r}
map <- get_googlemap("Sacramento, California", zoom = 11, maptype = "roadmap")
```

Using ggmap, I map the layers along with the addresses!
Woo hoo! 

```{r fig.width=8, fig.height=10,  fig.fullwidth=TRUE, message=FALSE, warning=FALSE}
map %>% ggmap() + 
    geom_sf(data = cc[1], 
            aes(fill = Name),
             inherit.aes = FALSE, alpha=0.5)  +
    geom_point(data = addresses, aes(x = long, y = lat), color = "navy", size = 1)
```

